---
title: "Processing SLCMA results"
output: html_document
---
This script reads in the SLCMA results for the four different analyses of 7 adversities and smoking exposure, combining them into usable formats for downstream analyses. The two resulting files are: 
- age7.all.results: large dataframe for 7 adversities x 4 analyses
- smoking.slcma: large dataframe for smoking x 4 analyses

Required files:
- ewas.results: obtained from running DD_EWAS_2021.Rmd

*Libraries and path*
```{r}
library(dplyr)
library(limma)

setwd("~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/")
load("/Users/alexlussier/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/1.Clean_Aug2021/Results/ewas_results_overlap_only_2021-08-21.Rdata", verbose=T)

cpgs.overlap <- unique(ewas.results$cpg)
length(cpgs.overlap)
rm(ewas.results)
```

*Analysis 1 - overlapping samples only (no twins)*
```{r}
path ="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/"

for(i in c("abuse","Fscore","mompsy","nbhqual","oneadult","parc", "r_faminst")){
  if(i=='abuse'){analysis.1 <- data.frame()}
  print(i)
  
  file <- list.files(path)[grep(i, list.files(path))]
  if(i == "mompsy"){i = "mompsych"}
  if(i == "parc"){i = "parcruelty"}
  if(i == "r_faminst"){i ="r_faminstability"}
  if(i == "Fscore"){
    file ="SLCMA_Fscore_fixed/01-oldDNAm-origCovars-covTest-Fscore_fixed_20210821.Rdata"
    }
  
  load(paste(path, file, sep=""))
  head(res)
  res <- res[,c("probe","Variable_Name","Improvement_R2","P-value")]
  res <- as.data.frame(lapply(res, unlist))
  colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
  res$Variable_Name <- gsub("fixed_","",res$Variable_Name)
  res <- res[which(res$CpG %in% cpgs.overlap),]
  res$FDR <- p.adjust(res$P.value, method = "BH")
  res$bonf <- p.adjust(res$P.value, method = "bonferroni")
  res$Model <- gsub("D", "", gsub(paste(i,"_", sep =""), "", res$Variable_Name))
  models <- unique(res$Model)
  models.2 <- models[-which(models %in% c("accumulation","ever","recency"))]
  months <- models.2[grep("m", models.2)][order(as.numeric(gsub("m", "", 
                                                                models.2[grep("m", models.2)]  )))]
  years <- models.2[grep("y", models.2)][order(as.numeric(gsub("y", "", 
                                                               models.2[grep("y", models.2)]  )))]
  models.cor <- c(months, years, models[-which(models %in% models.2)])
  res$Model <- factor(res$Model, levels = models.cor)
  
  res$Timing <- as.numeric(gsub("y", "", gsub("m","", res$Model)))
  res$Timing[which(res$Model %in% c("accumulation", "ever","recency"))]<-0
  res$Timing[grep("y", res$Model)] <- res$Timing[grep("y", res$Model)]*12
  res$Timing <- as.numeric(res$Timing)
  res$Timing[which(res$Timing <=36)]<-"very_early"
  res$Timing[which(res$Timing >36 & res$Timing<=61)]<-"early"
  res$Timing[which(res$Timing >61 & res$Timing <=84)]<-"middle"
  res$Timing[which(res$Timing %in% c(96, 108, 120,132))]<-"late"
  res$Timing[which(res$Model =="accumulation")]<-"accumulation"
  #res$Timing[which(res$Model =="ever")]<-"ever"
  res$Timing[which(res$Model =="recency")]<-"recency"
  res$Timing <- factor(res$Timing, levels = c("very_early","early","middle",
                                                          "accumulation","recency"))
  res$Adversity <- i
  analysis.1 <- rbind(analysis.1, res)
  rm(res, file, i, months, models, models.cor, models.2, years)
}

analysis.1$Method <- "covTest"
analysis.1$Covars <- "Original"
analysis.1$Analysis <- "1"
analysis.1$Data <- "Old"
analysis.1$Data <- factor(analysis.1$Data, levels = c("Old","New"))
analysis.1$unique <- paste(analysis.1$CpG, analysis.1$Adversity, analysis.1$Timing, sep="_")
analysis.1$unique_adversity <- paste(analysis.1$CpG, analysis.1$Adversity, sep="_")
analysis.1$unique_timing <- paste(analysis.1$CpG, analysis.1$Timing, sep="_")
head(analysis.1)
dim(analysis.1)

rm(path)

```

*Analysis 2*
```{r}
path ="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/data/SLCMA-results/02-newDNAm-origCovars-covTest/"

for(i in c("abuse","Fscore","mompsy","nbhqual","oneadult","parc", "r_faminst")){
  if(i=='abuse'){analysis.2 <- data.frame()}
  print(i)
  
  file <- list.files(path)[grep(i, list.files(path))]
  if(i == "mompsy"){i = "mompsych"}
  if(i == "parc"){i = "parcruelty"}
  if(i == "r_faminst"){i ="r_faminstability"}
  if(i == "Fscore"){
    load("~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA_Fscore_fixed/02-newDNAm-origCovars-covTest-Fscore_fixed_20210821.Rdata")
    }
    else{
    load(paste(path, file, sep=""))}
  
  res <- res[,c("probe","Variable_Name","Improvement_R2","P-value")]
  res <- as.data.frame(lapply(res, unlist))
  colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
  res$Variable_Name <- gsub("fixed_","",res$Variable_Name)
  res <- res[which(res$CpG %in% cpgs.overlap),]
  res$FDR <- p.adjust(res$P.value, method = "BH")
  res$bonf <- p.adjust(res$P.value, method = "bonferroni")
  res$Model <- gsub("D", "", gsub(paste(i,"_", sep =""), "", res$Variable_Name))
  models <- unique(res$Model)
  models.2 <- models[-which(models %in% c("accumulation","ever","recency"))]
  months <- models.2[grep("m", models.2)][order(as.numeric(gsub("m", "", 
                                                                models.2[grep("m", models.2)]  )))]
  years <- models.2[grep("y", models.2)][order(as.numeric(gsub("y", "", 
                                                               models.2[grep("y", models.2)]  )))]
  models.cor <- c(months, years, models[-which(models %in% models.2)])
  res$Model <- factor(res$Model, levels = models.cor)
  
  res$Timing <- as.numeric(gsub("y", "", gsub("m","", res$Model)))
  res$Timing[which(res$Model %in% c("accumulation", "ever","recency"))]<-0
  res$Timing[grep("y", res$Model)] <- res$Timing[grep("y", res$Model)]*12
  res$Timing <- as.numeric(res$Timing)
  res$Timing[which(res$Timing <=36)]<-"very_early"
  res$Timing[which(res$Timing >36 & res$Timing<=61)]<-"early"
  res$Timing[which(res$Timing >61 & res$Timing <=84)]<-"middle"
  res$Timing[which(res$Timing %in% c(96, 108, 120,132))]<-"late"
  res$Timing[which(res$Model =="accumulation")]<-"accumulation"
  #res$Timing[which(res$Model =="ever")]<-"ever"
  res$Timing[which(res$Model =="recency")]<-"recency"
  res$Timing <- factor(res$Timing, levels = c("very_early","early","middle",
                                                          "accumulation","recency"))
  res$Adversity <- i
  analysis.2 <- rbind(analysis.2, res)
  rm(res, file, i, months, models, models.cor, models.2, years)
}

analysis.2$Method <- "covTest"
analysis.2$Covars <- "Original"
analysis.2$Analysis <- "2"
analysis.2$Data <- "New"
analysis.2$Data <- factor(analysis.2$Data, levels = c("Old","New"))
analysis.2$unique <- paste(analysis.2$CpG, analysis.2$Adversity, analysis.2$Timing, sep="_")
analysis.2$unique_adversity <- paste(analysis.2$CpG, analysis.2$Adversity, sep="_")
analysis.2$unique_timing <- paste(analysis.2$CpG, analysis.2$Timing, sep="_")
head(analysis.2)
dim(analysis.2)

rm(path)
```

*Analysis 4*
```{r}
path ="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/data/SLCMA-results/04-newDNAm-origCovars-sI/"

for(i in c("abuse","Fscore","mompsy","nbhqual","oneadult","parc", "r_faminst")){
  if(i=='abuse'){analysis.4 <- data.frame()}
  print(i)
  
  file <- list.files(path)[grep(i, list.files(path))]
  if(i == "mompsy"){i = "mompsych"}
  if(i == "parc"){i = "parcruelty"}
  if(i == "r_faminst"){i ="r_faminstability"}
  if(i == "Fscore"){
    load("~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA_Fscore_fixed/04-newDNAm-origCovars-sI-Fscore_fixed_20210821.Rdata")
    }
    else{
    load(paste(path, file, sep=""))
      }
  
  res <- res[,c("probe","Variable_Name","R2","P-value")]
  res <- as.data.frame(lapply(res, unlist))
  colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
  res$Variable_Name <- gsub("fixed_","",res$Variable_Name)
  res <- res[which(res$CpG %in% cpgs.overlap),]
  res$FDR <- p.adjust(res$P.value, method = "BH")
  res$bonf <- p.adjust(res$P.value, method = "bonferroni")
  res$Model <- gsub("D", "", gsub(paste(i,"_", sep =""), "", res$Variable_Name))
  models <- unique(res$Model)
  models.2 <- models[-which(models %in% c("accumulation","ever","recency"))]
  months <- models.2[grep("m", models.2)][order(as.numeric(gsub("m", "", 
                                                                models.2[grep("m", models.2)]  )))]
  years <- models.2[grep("y", models.2)][order(as.numeric(gsub("y", "", 
                                                               models.2[grep("y", models.2)]  )))]
  models.cor <- c(months, years, models[-which(models %in% models.2)])
  res$Model <- factor(res$Model, levels = models.cor)
  
  res$Timing <- as.numeric(gsub("y", "", gsub("m","", res$Model)))
  res$Timing[which(res$Model %in% c("accumulation", "ever","recency"))]<-0
  res$Timing[grep("y", res$Model)] <- res$Timing[grep("y", res$Model)]*12
  res$Timing <- as.numeric(res$Timing)
  res$Timing[which(res$Timing <=36)]<-"very_early"
  res$Timing[which(res$Timing >36 & res$Timing<=61)]<-"early"
  res$Timing[which(res$Timing >61 & res$Timing <=84)]<-"middle"
  res$Timing[which(res$Timing %in% c(96, 108, 120,132))]<-"late"
  res$Timing[which(res$Model =="accumulation")]<-"accumulation"
  #res$Timing[which(res$Model =="ever")]<-"ever"
  res$Timing[which(res$Model =="recency")]<-"recency"
  res$Timing <- factor(res$Timing, levels = c("very_early","early","middle",
                                                          "accumulation","recency"))
  res$Adversity <- i
  analysis.4 <- rbind(analysis.4, res)
  rm(res, file, i, months, models, models.cor, models.2, years)
}
analysis.4$Method <- "sI"
analysis.4$Covars <- "Original"
analysis.4$Analysis <- "4"
analysis.4$Data <- "New"
analysis.4$Data <- factor(analysis.4$Data, levels = c("Old","New"))
analysis.4$unique <- paste(analysis.4$CpG, analysis.4$Adversity, analysis.4$Timing, sep="_")
analysis.4$unique_adversity <- paste(analysis.4$CpG, analysis.4$Adversity, sep="_")
analysis.4$unique_timing <- paste(analysis.4$CpG, analysis.4$Timing, sep="_")
head(analysis.4)
dim(analysis.4)

rm(path)
```

*Analysis 6*
```{r}
path ="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/data/SLCMA-results/06-newDNAm-educFWL-sI/"

for(i in c("abuse","Fscore","mompsy","nbhqual","oneadult","parc", "r_faminst")){
  if(i=='abuse'){analysis.6 <- data.frame()}
  print(i)
  
  file <- list.files(path)[grep(i, list.files(path))]
  if(i == "mompsy"){i = "mompsych"}
  if(i == "parc"){i = "parcruelty"}
  if(i == "r_faminst"){i ="r_faminstability"}
  if(i == "Fscore"){
    load("~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA_Fscore_fixed/06-newDNAm-educFWL-sI-Fscore_fixed_20210821.Rdata")
    }
    else{
    load(paste(path, file, sep=""))
      }
  
  res <- res[,c("probe","Variable_Name","R2","P-value")]
  res <- as.data.frame(lapply(res, unlist))
  colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
  res$Variable_Name <- gsub("fixed_","",res$Variable_Name)
  res <- res[which(res$CpG %in% cpgs.overlap),]
  res$FDR <- p.adjust(res$P.value, method = "BH")
  res$bonf <- p.adjust(res$P.value, method = "bonferroni")
  res$Model <- gsub("D", "", gsub(paste(i,"_", sep =""), "", res$Variable_Name))
  models <- unique(res$Model)
  models.2 <- models[-which(models %in% c("accumulation","ever","recency"))]
  months <- models.2[grep("m", models.2)][order(as.numeric(gsub("m", "", 
                                                                models.2[grep("m", models.2)]  )))]
  years <- models.2[grep("y", models.2)][order(as.numeric(gsub("y", "", 
                                                               models.2[grep("y", models.2)]  )))]
  models.cor <- c(months, years, models[-which(models %in% models.2)])
  res$Model <- factor(res$Model, levels = models.cor)
  
  res$Timing <- as.numeric(gsub("y", "", gsub("m","", res$Model)))
  res$Timing[which(res$Model %in% c("accumulation", "ever","recency"))]<-0
  res$Timing[grep("y", res$Model)] <- res$Timing[grep("y", res$Model)]*12
  res$Timing <- as.numeric(res$Timing)
  res$Timing[which(res$Timing <=36)]<-"very_early"
  res$Timing[which(res$Timing >36 & res$Timing<=61)]<-"early"
  res$Timing[which(res$Timing >61 & res$Timing <=84)]<-"middle"
  res$Timing[which(res$Timing %in% c(96, 108, 120,132))]<-"late"
  res$Timing[which(res$Model =="accumulation")]<-"accumulation"
  #res$Timing[which(res$Model =="ever")]<-"ever"
  res$Timing[which(res$Model =="recency")]<-"recency"
  res$Timing <- factor(res$Timing, levels = c("very_early","early","middle",
                                                          "accumulation","recency"))
  res$Adversity <- i
  analysis.6 <- rbind(analysis.6, res)
  rm(res, file, i, months, models, models.cor, models.2, years)
}

analysis.6$Method <- "sI"
analysis.6$Covars <- "Updated"
analysis.6$Analysis <- "6"
analysis.6$Data <- "New"
analysis.6$Data <- factor(analysis.6$Data, levels = c("Old","New"))
analysis.6$unique <- paste(analysis.6$CpG, analysis.6$Adversity, analysis.6$Timing, sep="_")
analysis.6$unique_adversity <- paste(analysis.6$CpG, analysis.6$Adversity, sep="_")
analysis.6$unique_timing <- paste(analysis.6$CpG, analysis.6$Timing, sep="_")
head(analysis.6)
dim(analysis.6)

rm(path)
```

*Combining adversity results*
```{r}
age7.all.results <- rbind(analysis.1,
                          analysis.2,
                          analysis.4,
                          analysis.6)
dim(age7.all.results)/4/7
head(age7.all.results)
save(age7.all.results, file="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/1.Clean_Aug2021/Results/age7.all.results.2021-09-21.Rdata")

levels(factor(age7.all.results$Timing))

rm(analysis.1, analysis.2, analysis.4, analysis.6, age7.all.results)
```

*Loading smoking SLCMA results*
```{r}
load("/Users/alexlussier/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA-oldDNAm-rerun-covTest-smoking_20200825.rdata", verbose=T)
head(res)
res <- res[,c("probe","Variable_Name","Improvement_R2","P-value")]
res <- as.data.frame(lapply(res, unlist))
colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
res <- res[which(res$CpG %in% cpgs.overlap),]
res$FDR <- p.adjust(res$P.value, method = "BH")
res$bonf <- p.adjust(res$P.value, method = "bonferroni")
res$Model <- gsub("smoke.trim.", "", res$Variable_Name)
res$Model <- factor(res$Model, levels = c("first","second","third","accumulation","recency"))
res$Timing <- res$Model
res$Adversity <- "Smoking"
res$Method <- "CovTest"
res$Covars <- "Original"
res$Analysis <- 1
res$Data <- "Old"
res$unique <- paste(res$CpG, res$Adversity, res$Timing, sep ="_")
res$unique_adversity <- paste(res$CpG, res$Adversity, sep ="_")
res$unique_timing <- paste(res$CpG, res$Timing, sep ="_")
smoking.slcma <- res
rm(res)

#new data, old analysis
load("/Users/alexlussier/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA-newDNAm-rerun-covTest-smoking_20200825.rdata", verbose=T)
dim(res)
res <- res[,c("probe","Variable_Name","Improvement_R2","P-value")]
res <- as.data.frame(lapply(res, unlist))
colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
res <- res[which(res$CpG %in% cpgs.overlap),]
res$FDR <- p.adjust(res$P.value, method = "BH")
res$bonf <- p.adjust(res$P.value, method = "bonferroni")
res$Model <- gsub("smoke.trim.", "", res$Variable_Name)
res$Model <- factor(res$Model, levels = c("first","second","third","accumulation","recency"))
res$Timing <- res$Model
res$Adversity <- "Smoking"
res$Method <- "CovTest"
res$Covars <- "Original"
res$Analysis <- 2
res$Data <- "New"
res$unique <- paste(res$CpG, res$Adversity, res$Timing, sep ="_")
res$unique_adversity <- paste(res$CpG, res$Adversity, sep ="_")
res$unique_timing <- paste(res$CpG, res$Timing, sep ="_")
smoking.slcma <- rbind(smoking.slcma, res)
rm(res)

#new dataset results - updated analysis, original covariates
load("/Users/alexlussier/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA-newDNAm-rerun-sI-smoking_20200825.rdata", verbose=T)
res <- res[,c("probe","Variable_Name","R2","P-value")]
res <- as.data.frame(lapply(res, unlist))
colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
res <- res[which(res$CpG %in% cpgs.overlap),]
res$FDR <- p.adjust(res$P.value, method = "BH")
res$bonf <- p.adjust(res$P.value, method = "bonferroni")
res$Model <- gsub("smoke.trim.", "", res$Variable_Name)
res$Model <- factor(res$Model, levels = c("first","second","third","accumulation","recency"))
res$Timing <- res$Model
res$Adversity <- "Smoking"
res$Method <- "sI"
res$Covars <- "Original"
res$Analysis <- 4
res$Data <- "New"
res$unique <- paste(res$CpG, res$Adversity, res$Timing, sep ="_")
res$unique_adversity <- paste(res$CpG, res$Adversity, sep ="_")
res$unique_timing <- paste(res$CpG, res$Timing, sep ="_")
smoking.slcma <- rbind(smoking.slcma, res)
rm(res)

#new dataset results - updated analysis
load("/Users/alexlussier/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/results/SLCMA-newDNAm-rerun-sI-educFWL-smoking_20200805.Rdata", verbose=T)
res <- res[,c("probe","Variable_Name","R2","P-value")]
res <- as.data.frame(lapply(res, unlist))
colnames(res) <- c("CpG", "Variable_Name", "R2","P.value")
res <- res[which(res$CpG %in% cpgs.overlap),]
res$FDR <- p.adjust(res$P.value, method = "BH")
res$bonf <- p.adjust(res$P.value, method = "bonferroni")
res$Model <- gsub("smoke.trim.", "", res$Variable_Name)
res$Model <- factor(res$Model, levels = c("first","second","third","accumulation","recency"))
res$Timing <- res$Model
res$Adversity <- "Smoking"
res$Method <- "sI"
res$Covars <- "Updated"
res$Analysis <- 6
res$Data <- "New"
res$unique <- paste(res$CpG, res$Adversity, res$Timing, sep ="_")
res$unique_adversity <- paste(res$CpG, res$Adversity, sep ="_")
res$unique_timing <- paste(res$CpG, res$Timing, sep ="_")
smoking.slcma <- rbind(smoking.slcma, res)
rm(res)

dim(smoking.slcma)/4
identical(colnames(smoking.slcma), colnames(age7.all.results))#TRUE

save(smoking.slcma, file="~/Dropbox (Partners HealthCare)/Dunn_shared/ALSPAC/Epigenetic/Paper Draft - Data differences at age 7/DNAm-data-diff-analysis/1.Clean_Aug2021/Results/smoking.slcma.results.2021-09-21.Rdata")

rm(smoking.slcma)
```
